---
title: "Credit Curve bootstrapping"
author: "Bertrand Le Nezet"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style


$1 + 1$

$D_{KL}(P\Vert Q)=\int_{-\infty}^{\infty}p(x)\ln\frac{p(x)}{q(x)}dx$

$$ Y = \beta_0 + \beta_1 x + \epsilon$$

## Credit Curve Bootstrapping

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

RadioShack Corp, on 27-May-2014
Pfizer Inc


```{r, fig.show='hold'}
library(credule)

yieldcurveTenor = c(1,2,3,4,5,7)
yieldcurveRate = c(0.0050,0.0070,0.0080,0.0100, 0.0120,0.0150)
cdsTenor = c(1,2,3,4,5,7,10,15,20,30)
cdsSpread = c(0.6405,0.5956,0.5511,0.5144,0.4894,0.4511,0.4156,0.3815,0.3657,0.3506)
premiumFrequency = 4
defaultFrequency = 12
accruedPremium = TRUE
RR = 0.40

res = bootstrapCDS(yieldcurveTenor,
             yieldcurveRate,
             cdsTenor,
             cdsSpread,
             RR,
             premiumFrequency,
             defaultFrequency,
             accruedPremium)

cdsQuote = data.frame(t(cdsSpread))
colnames(cdsQuote) = cdsTenor
knitr::kable(cdsQuote)


default.par = par()
par(mfrow=c(1,2),oma = c(0, 0, 2, 0))
barplot(res$hazrate, names.arg = res$tenor, xlab="Time", ylab = "Non-Cummulative Hazard rate", main ="Hazard Rate (Non-Cum)")
plot(res$tenor, res$survprob, type = "l", xlab="Time", ylab = "Survival Probability", main ="Survival Probability" )
mtext("RadioShack Corp on 27-May-2014", outer = TRUE, cex = 1.5)
par = default.par

cdsSpread = c(0.0003,0.0009,0.0015,0.0021,0.0028,0.0043,0.0061,0.0063,	0.0068,0.0066)

res = bootstrapCDS(yieldcurveTenor,
             yieldcurveRate,
             cdsTenor,
             cdsSpread,
             RR,
             premiumFrequency,
             defaultFrequency,
             accruedPremium)

barplot(res$hazrate)
plot(res$tenor, res$survprob, type = "l")

```



64.05%	59.56%	55.11%	51.44%	48.94%	45.11%	41.56%	38.15%	36.57%	35.06%	27.44%


## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
